---
title: "Activity Monitor"
author: "Anand Mohan"
date: "Sunday, September 13, 2015"
output: html_document
---

Loading and preprocessing the data

``` {r load_data}
loadFile <- read.csv("Activity.csv")
```

The following code piece will output the total number of steps taken per day
```{r activity_mon}
file2 <- aggregate(steps~date, data=loadFile, FUN=sum)
head(file2)
```


The next part of the assignment is to plot a histogram of the total number of steps taken per day:
```{r activity_mon_hist}
hist(file2$steps, main="steps taken per day", border="blue", col="grey", xlim=c(1000,30000), las=1, breaks=5, xlab="Steps")
```

The mean of the total number of steps taken per day:
```{r activity_mon_mean}
dfMean <- aggregate(steps~date, data=loadFile, FUN=mean)
head(dfMean)
```

The median of the total number of steps taken per day:
```{r activity_mon_med}
dfMedian <- aggregate(steps~date, data=loadFile, FUN=median)
head(dfMedian)
```

Answers to the daily activity pattern..

Time series plot of  5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r activity_mon_mean_across_alldays}
stepsMean <- aggregate(steps~interval, data=loadFile, FUN=mean)
stepsMeants <- ts(stepsMean$steps)
plot.ts(stepsMeants)
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r activity_mon_max}
maxValIndex <- which.max(stepsMean$steps)
print(maxValIndex)
```

The number of NAs in the dataset is 
```{r num_of_NAs}
print(sum(is.na(loadFile$steps)))
```

Now, we will create a new dataset with all the missing values replaced with the mean of steps for that day.
```{r newDataFile}
library(Hmisc)
newActFile <- loadFile
newActFile$steps <- impute(loadFile$steps, fun=mean)
```

Total Number of steps calculated with the new datafile is
```{r newDataFile_total_steps}
newdfSum <- aggregate(steps~date, data=newActFile, FUN=sum)
head(newdfSum)
```

Histogram of the total number of steps taken per day with the new file:
```{r newFile_hist}
hist(newdfSum$steps, main="steps taken per day", border="blue", col="grey", xlim=c(1000,30000), las=1, breaks=5, xlab="Steps")
```

The mean of the total number of steps taken per day with new file:
```{r newFile_mean}
newdfMean <- aggregate(steps~date, data=newActFile, FUN=mean)
head(newdfMean)
```

The median of the total number of steps taken per day with new file:
```{r newFile_med}
newdfMedian <- aggregate(steps~date, data=newActFile, FUN=median)
head(newdfMedian)
```

Create a new file with a new column marking weekdays and weekends
```{r newFile_dayType}
newActFile$day <-  ifelse(as.POSIXlt(newActFile$date)$wday %in% c(0,6), 'weekend', 'weekday')
head(newActFile)
```

panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)
``` {r panelplot}
newStepsMean <- aggregate(steps~interval+day, data=newActFile, FUN=mean)
ggplot(newStepsMean, aes(interval, steps)) + 
        geom_line() + 
        facet_grid(day ~ .) +
        xlab("interval") + 
        ylab("steps mean")
```
